---
title: "MSDS6306 Case Study 1"
author: "Ivan Charkin, Laura Niederlander, Shon Mohsin"
date: "June 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


We have been asked by the product development group at National Beverage Corporation to review the data for RelaxBev which they recently accquired. We will assess the number of breweries, the median alcohol levels, and which state has the highest ABV/IBU levels. Based on this information, the product development team hope to make some decisions about the integration of these breweries and beers into their current product offerings. Additionally, it will provide information about the current number of breweries to assess any overlap with their own facilities. 


###1. How many breweries are present in each state?
```{r}
#Load data from Github
br <- read.csv("https://raw.githubusercontent.com/icharkin/ICharkin_MSDS6306_HW/master/CaseStudy1/Breweries.csv", header=TRUE)
beers <- read.csv ("https://raw.githubusercontent.com/icharkin/ICharkin_MSDS6306_HW/master/CaseStudy1/Beers.csv", header=TRUE)

brnum<-summary (br$State)
brnum
```
We have assessed that the current number of breweres span across most states with Colorado and California having the most facilities at 47 and 39 respectfully.  Based on this list, there are states where National Beverage Corporation currently does not have coverage.

This includes: South Dakota, Washington DC, and Utah. These facilities could help to store product for distribution as well as backup  product manufacturing if necessary. 



###2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
Observations:
We have provided a snapshot of the data for your reference below. As you can see, the data included the Brewery Name, City, State, Name of the beer produced there, Beer ID, ABV and IBU. 

RelaxBev uses AVB (Alcohol by Volume of Beer) and IBU (International Bitterness Units of beer) to determine the formula levels for optimal taste. The measurements are key to reproducting these products to the specification to reach the desireable outcome.


```{r}
#Align the names between datasets
#Only columns common to the two datasets is Brewery ID 
colnames(beers)[1] <- "Beer_name"
colnames(br)[1] <- "Brewery_id"
colnames(br)[2] <- "Brewery_name"

merged <- merge(br, beers)
head (merged)
```

The data are successfully merged. 

###3. Report the number of NA's in each column.

During our analysis process, we found that several fields were missing data. The total number of fields with incomplete data included ABV with 62 fields, and IBU with 1005 fields. Without these fields, it would be difficult to produce the beers associated with these levels. We would recommend pursuing RelaxBev to complete this dataset.



```{r}
colSums(is.na(merged))
```
All NA's are in ABV and IBU columns. We have to be careful with the further analysis of these variables as it is obvious that the data set is incomplete. 

###4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

The chart below provides a histogram of the ABV and IBUs at the Median point by State. This overview by State shows that West Virginia has the highest median at 0.06. The lowest median for ABV level is in Utah at 0.04.  South Dakota has the lowest IBUs median with 0. The States with the highest median in IBU are Maine and West Virginia at 60 parts per million.


```{r}
ABV<-tapply (merged$ABV, merged$State, median, na.rm = TRUE)
barplot(ABV, las=2, cex.names=0.7)

IBU<-tapply (merged$IBU, merged$State, median, na.rm = TRUE)
barplot(IBU, las=2, cex.names=0.7)
```

There is much more variation in the Bitterness between states than in alcohol content. 

###5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

This overview by State shows that Colorado has beer with the highest ABV. The State with the beer with the highest IBU is Oregon. 

```{r}
as.character (merged$State [which (merged$ABV==max (merged$ABV, na.rm=TRUE) )]  )
as.character (merged$State [which (merged$IBU==max (merged$IBU, na.rm=TRUE) )]  )
```
Colorado has the beer with the maximum alcohol content. 
OR has the most bitter beer. 

###6. Summary statistics for the ABV variable.
```{r}
summary(merged$ABV)
```


###7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
The scatter plot below does show evidence of a relationship between bitterness and alcohol content. As the bitterness units increases the alcohol content increases as well. There is statistically significant relationship between these two variables (P<2e-16). Based on the data on average for every increasing IBU unit we should expect 1.04e-5 increase in alcohol content. 

```{r}
plot(merged$IBU, merged$ABV)
abline(lm(merged$ABV ~ merged$IBU))
summary (lm(merged$ABV ~ merged$IBU))
```
